---
title: "Quick analysis"
author: "Nicole Greco"
format: html
editor: visual
---

# Introduction

## About the data

The data for this analysis were downloaded from the [Arctic Data Center](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X) on 10/7/2024. 

## Setup

```{r}
library(readr)
library(here)

```


## Read in the data

```{r}
#| warning: false
bg_chem<-read_csv(here('data','BGchem2008data.csv'))

```

# Analysis

## Calculate summary statistics

```{r}
#| eval: false
#| echo: false
## print the column names
colnames(bg_chem)

## get the general structure
str(bg_chem)

## first six lines
head(bg_chem)

## get a summary of each column
summary(bg_chem)

## unique values of a column
unique(bg_chem$Date)
```

calculate tje summary statistics (means) for nitrate, nitrite, ammonia, and phosphate
```{r}
nitrate<-mean(bg_chem$NO3)
nitrite<-mean(bg_chem$NO2)
amm<-mean(bg_chem$NH4)
phos<-mean(bg_chem$P)

```

## Calculate the mean Redfield ratio
Redfield ratio is the relative proportion of nitrogen to phosphorus
```{r}
ratio<-(nitrate+nitrite+amm)/phos

```

the Redfield ratio for this dataset is approximately `r round(ratio)`

## Plot Redfield ratio

```{r}
#| echo: false
plot(bg_chem$P,bg_chem$NO2+bg_chem$NO3+bg_chem$NH4)

```


# Conclusion

